---
title: "covid 19"
author: "Trevor Ragazzo"
date: "7/13/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```


```{r eval=FALSE}
kRdata = kRdata %>%
mutate(across(where(is.character), as.factor))
```


### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$ There is no correlation between sex and dying 

$H_A$ There is a correlation between sex and dying


Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
dd2 <- kRdata %>% group_by(Sex,result) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill= result))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This GGplot shows that we see more females than males that were observed. We see that more males seemed to die than females. We saw less males being observed and about 65% of males died.

  
## Numerical results


```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

From this graph we see that double the amount of males have died compared to females. Almost 3% males died and 1.4% females died. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In this test we see that men are about 2.3 times more likely to die than women. The P-value for getting these results are .0012. 


## Conclusions

First the barchart shows us results of about 65% of males die but only 35% of females die. As we dig deeper into this we see the numerical results showing that men are 2.3 times more likely to die rather than a female. Lastly we look into the P-Value and see that the chances of this changing is about .0012. 